---
title: "Evaluation of Marginal Likelihood"
author: "Philip Cooney"
date: "03 April 2020"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
link-citations: yes
bibliography:
- References_Gibs.bib

---


```{r, include = FALSE}

setwd("D:/PhD offline/")
source("C:/Users/phili/OneDrive/PhD/R codes/Functions Gibbs Markdown.R")
source("Functions Reversible Jump.R")
library("Brobdingnag")

```
In a previous section we discussed how the marginal likelihood of a piecewie exponential data segment is: 

$$\frac{\Gamma(\alpha+n)}{\Gamma(\alpha)}\frac{\beta^\alpha}{(\beta+ \sum_{i = 1}^n t_i)^{\alpha+n}}$$

Computational issues may arise for larger values of $n$ and $\sum t_i$ both the numerator and denominator will be very large numbers.

For simplicity we assume $\alpha$ and $\beta$ are 1 which means the above equation can be rewritten as (note $\Gamma(1)= 1$):
$$\frac{\Gamma(\alpha+n)}{(\beta+ \sum_{i = 1}^n t_i)^{\alpha+n}}$$
Taking the logarithm of this expression we have

$$\text{log}\bigg[\frac{\Gamma(\alpha+n)}{(\beta+ \sum_{i = 1}^n t_i)^{\alpha+n}}\bigg]$$
which can be expressed as:

$$\text{Ln }\Gamma(\alpha+n) -(\alpha+n)\text{log}(\beta+ \sum_{i = 1}^n t_i)$$ where $\text{Ln }\Gamma$ is the Log Gamma Function
We have obtained the log of the probability which will be useful when we wish to calculate the ratio of the two probabilities for the reversible jump algorithm.

$$ \frac{\text{Prob Propose}}{\text{Prob current}}= \frac{e^{\text{Log Prob Propose}}}{e^{\text{Log Prob Current}}} = e^{\text{Log Prob Propose-Log Prob Current}}$$
which follows from the properties of exponents.

We can use the integrate function in R to verify our results by writing a density function for the marginal likelihood expression $\frac{\Gamma(\alpha+n)}{\Gamma(\alpha)}\frac{\beta^\alpha}{(\beta+ \sum_{i = 1}^n t_i)^{\alpha+n}}$ and evaluate for $n = 1$ and $\sum t = 10$ with $\alpha, \beta = 1$

```{r}


dmarg.gamma <- function(x,n, time, alpha.hyper =1,beta.hyper =1){
  dens<-((beta.hyper^alpha.hyper)/gamma(alpha.hyper))*(x^(alpha.hyper+n -1))*
    exp(-(time+beta.hyper)*x)
return(dens)
}


integrate(dmarg.gamma,n = 1, time = 10,
          lower = 0,upper = Inf ) 
exp(marg.lik.eval.log(cbind(1,10),1,1))

```

The equality of the results verifies the formula.






